Positional entropy, conditional entropy and mutual information in English CHILDES corpora

### 2019-04-24 surprisal/childes_surprisal.Rmd
trigram providence surprisals, second position not recording

```{r trigrams}
adult_trigrams <- adult_utterances %>%
  group_by(utterance_id) %>%
  mutate(lag_word1 = lag(word)) %>%
  mutate(lag_word2 = lag(lag_word1)) %>%
  group_by(lag_word2, lag_word1, word) %>%
  count() %>%
  filter(!is.na(lag_word1), !is.na(lag_word2)) %>%
  ungroup() %>%
  mutate(tri_joint_p = n / sum(n)) %>%
  select(-n) %>%
  left_join(adult_bigrams, by = c("lag_word2" = "lag_word", "lag_word1" = "word")) %>%
  mutate(tri_cond_p = tri_joint_p / joint_p) %>%
  select(-cond_p, -joint_p)

trigram_surprisals_prep <- adult_utterances %>%
  group_by(utterance_id) %>%
  mutate(lag_word1 = lag(word)) %>%
  mutate(lag_word2 = lag(lag_word1)) %>%
  left_join(adult_trigrams) %>%
  left_join(adult_bigrams) %>%
  left_join(adult_unigrams)

trigram_surprisals <- trigram_surprisals_prep %>%
  filter(!is.na(lag_word)) %>%
  mutate(s = -log(cond_p)) %>%
  group_by(length, word_order) %>%
  summarise(s = mean(s)) %>%
  filter(word_order <= length)

trigram_surprisals_nonnull <- trigram_surprisals_prep %>%
  mutate(s = ifelse(is.na(lag_word), -log(p), -log(cond_p))) %>%
  group_by(length, word_order) %>%
  summarise(s = mean(s)) %>% 
  filter(word_order <= length)

>>>>

```{r trigrams}
adult_trigrams <- adult_utterances %>%
  group_by(utterance_id) %>%
  mutate(lag_word1 = lag(word)) %>%
  mutate(lag_word2 = lag(lag_word1)) %>%
  group_by(lag_word2, lag_word1, word) %>%
  count() %>%
  filter(!is.na(lag_word1), !is.na(lag_word2)) %>%
  ungroup() %>%
  mutate(tri_joint_p = n / sum(n)) %>%
  select(-n) %>%
  left_join(adult_bigrams, by = c("lag_word2" = "lag_word", "lag_word1" = "word")) %>%
  mutate(tri_cond_p = tri_joint_p / joint_p) %>%
  select(-joint_p, -tri_joint_p)

trigram_surprisals_prep <- adult_utterances %>%
  group_by(utterance_id) %>%
  mutate(lag_word1 = lag(word)) %>%
  mutate(lag_word2 = lag(lag_word1)) %>%
  left_join(adult_trigrams) %>% #add bigram and trigram conditionals
  left_join(adult_unigrams)

trigram_surprisals_nonnull <- trigram_surprisals_prep %>%
  mutate(s = ifelse(is.na(lag_word1), # check if it's not the first word
                    ifelse(is.na(lag_word2), # check if not the second word
                           -log(p), # trigram if third or beyond
                           -log(cond_p)), # bigram if second word
                    -log(tri_cond_p))) %>% # unigram if first word
  group_by(length, word_order) %>%
  summarise(s = mean(s)) %>% 
  filter(word_order <= length)

trigram_surprisals_nonnull %>%
  filter(length %in% c(5, 7, 9, 11, 13, 15)) %>%
  ggplot(aes(x = word_order, y = s)) + 
  facet_wrap(~ length) + 
  geom_point() + 
  geom_smooth(se = F)
